Toward Resolved Simulations of Burning Fronts in Thermonuclear X-ray Bursts

### 2018-09-27 astronum_2018/astronum-xrb.tex
more cites

A longstanding goal is to perform simulations where we resolve the
flame structure, to allow us to accurately capture the
nucleosynthesis, and watch it move across the neutron star surface.
The results of these simulations will enable strict comparison to
observations made with the new generation of X-ray telescopes such as
\textit{eXTP}\/ and \textit{STROBE-X}\/
\cite{art-2017-wilhod-etal,art-2016-zhang-etal} which will offer
high collecting area and time resolution.
We show some in-progress calculations that are a step toward this, discuss
their remaining approximations, and also discuss what new algorithmic
techniques might be needed to realize this goal with the advent of
exascale computing.

@@

If we continue to resolve the flame structure, then we need algroithms
that can do so less expensively.  We can be smarter about our grid
refinement strategy, putting the finest grids only at the very base of
the burning layer, where the flame is the thinnest.  Using
well-balanced schemes can help us retain hydrostatic equilibrium (HSE)
in the lower refined regions, reducing the number of zones needed to
maintain fidelity.  We can also explore initializing newly created fine
grids ahead of the burning with an understanding of hydrostatic
equilibrium to allow us to only refine the peak burning as the flame
moves across the star.

@@

Other ways to make the simulation less expensive can include a subgrid
model, Flame models are commonly used for Type Ia supernovae.  There
however, the flame is thin compared to the scale height, so a simple
model can represent the unresolved flame on the grid.  For the XRB,
the flame thickness is a non-trivial fraction of the pressure scale
height and further, the laminar flame properties will change strongly
over the height of the burning layer (see~\cite{Timmes00}).  This
makes traditional flame models a poor fit.  Methods that use even
something as simple as subzone burning.  This is harder...

>>>>

A longstanding goal is to perform simulations where we resolve the
flame structure, to allow us to accurately capture the
nucleosynthesis, and watch it move across the neutron star surface.
The results of these simulations will enable strict comparison to
observations made with the new generation of X-ray telescopes such as
\textit{eXTP}\/ and \textit{STROBE-X}\/
\cite{art-2017-wilhod-etal,art-2016-zhang-etal,intZand2018} which will offer
high collecting area and time resolution.
We show some in-progress calculations that are a step toward this, discuss
their remaining approximations, and also discuss what new algorithmic
techniques might be needed to realize this goal with the advent of
exascale computing.

@@

If we continue to resolve the flame structure, then we need algroithms
that can do so less expensively.  We can be smarter about our grid
refinement strategy, putting the finest grids only at the very base of
the burning layer, where the flame is the thinnest.  Using
well-balanced schemes~\cite{kappeli:2016} can help us retain hydrostatic equilibrium (HSE)
in the lower refined regions, reducing the number of zones needed to
maintain fidelity.  We can also explore initializing newly created fine
grids ahead of the burning with an understanding of hydrostatic
equilibrium to allow us to only refine the peak burning as the flame
moves across the star.

@@

Other ways to make the simulation less expensive can include a subgrid
model, Flame models are commonly used for Type Ia supernovae.  There
however, the flame is thin compared to the scale height, so a simple
model can represent the unresolved flame on the grid.  For the XRB,
the flame thickness is a non-trivial fraction of the pressure scale
height and further, the laminar flame properties will change strongly
over the height of the burning layer (see~\cite{Timmes00}).  This
makes traditional flame models a poor fit.  Methods that use even
something as simple as subzone burning (e.g.~\cite{Wang2012190}).  This is harder...
